
Traffic Collision in NYC (2013-2018)

This analysis of traffic collisions in New York City focuses on factors that contribute to the likelihood of being killed or injured (hereafter referred to as a casualty) in an accident during the time period between years 2013 and 2018.

This analysis consists of seven parts:

The first part consists of data pre-processing and exploratory analysis. After this the total number of injured persons were calculated for the year 2018.
The second part addresses the question of what proportion of all collisions in the year 2016 occurred in Brooklyn.
The third part addresses the question of what proportion of collisions in the year 2016 resulted in injury or death of a cyclist.
The fourth part required computation of the number of accidents per capita involving alcohol in 2017 for each borough. Highest value was to be reported.
Fifth part required finding the zip code with the highest number of vehicles involved in collisions for the year 2018.
Sixth part was based on linear regression and reporting the slope of the regression line for total number of collisions vs the time period in years.
Finally, the seventh part involved validating the null hypothesis: The total number of collisions are ambivalent to a particular month for the year 2017. A chi squared test was performed.


This analysis is based on the NYPD Motor Vehicle Collisions dataset available here: https://data.cityofnewyork.us/Public-Safety/NYPD-Motor-Vehicle-Collisions/h9gi-nx95

The population data for different boroughs in NY was taken from here:
https://en.wikipedia.org/wiki/Boroughs_of_New_York_City
#Ans to q1
```{r}
data = read.csv("C:/Users/Malik/Desktop/NYPD_Motor_Vehicle_Collisions.csv")
library(dplyr)
options(digits = 10)
data$DATE <- as.Date(data$DATE, "%m/%d/%Y")
data=data[data$DATE <= "2018-12-31",]
TPI = sum(data$NUMBER.OF.PERSONS.INJURED[!is.na(data$NUMBER.OF.PERSONS.INJURED)])
TPI
```

```{r}
#splitting data for year 2016 only
dat2=data[data$DATE <= "2016-12-31" & data$DATE >= "2016-01-01",]
count_borough <- count(dat2, BOROUGH)
count_Brook = count_borough[3,2]
#removing null values as stated in q2
count_borough = count_borough[-1,]
prop_brook= count_Brook/sum(count_borough$n)
prop_brook
```

#Ans to q2


#Ans to q3
```{r}
count_nodi <- count(dat2,NUMBER.OF.CYCLIST.INJURED =="0" & NUMBER.OF.CYCLIST.KILLED =="0")
prop2 <- count_nodi[1,2]/sum(count_nodi$n)
prop2
```

```{r}
#loading data for q4 pop of different bouroghs
boropop = read.csv("C:/Users/Malik/Desktop/boropop.csv")
#splitting data for year 2017 only
data2017=data[data$DATE <= "2017-12-31" & data$DATE >= "2017-01-01",]

count_2017_BR = count(data2017,BOROUGH=="BRONX" & (CONTRIBUTING.FACTOR.VEHICLE.1=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.2=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.3=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.4=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.5=="Alcohol Involvement"))
count_2017_SI = count(data2017,BOROUGH=="STATEN ISLAND" & (CONTRIBUTING.FACTOR.VEHICLE.1=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.2=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.3=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.4=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.5=="Alcohol Involvement"))
count_2017_Queen = count(data2017,BOROUGH=="QUEENS" & (CONTRIBUTING.FACTOR.VEHICLE.1=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.2=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.3=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.4=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.5=="Alcohol Involvement"))
count_2017_BROK = count(data2017,BOROUGH=="BROOKLYN" & (CONTRIBUTING.FACTOR.VEHICLE.1=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.2=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.3=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.4=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.5=="Alcohol Involvement"))
count_2017_MAN <- count(data2017,BOROUGH=="MANHATTAN" & (CONTRIBUTING.FACTOR.VEHICLE.1=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.2=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.3=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.4=="Alcohol Involvement"| CONTRIBUTING.FACTOR.VEHICLE.5=="Alcohol Involvement"))

#finding prop with the help of pop count for each borough
BROK_CAP = count_2017_BROK[2,2]/boropop[2,2]
MAN_CAP  = count_2017_MAN[2,2]/boropop[3,2]
SI_CAP = count_2017_SI[2,2]/boropop[5,2]
BRONX_CAP = count_2017_BR[2,2]/boropop[1,2]
QUE_CAP = count_2017_Queen[2,2]/boropop[4,2]
max_cap = max(BROK_CAP,MAN_CAP,SI_CAP,BRONX_CAP,QUE_CAP)
max_cap
```

#Brooklyn had the highest prop . Ans Q4
```{r}
data2016<- data.frame(dat2$ZIP.CODE,dat2$CONTRIBUTING.FACTOR.VEHICLE.1,dat2$CONTRIBUTING.FACTOR.VEHICLE.2,dat2$CONTRIBUTING.FACTOR.VEHICLE.3,dat2$CONTRIBUTING.FACTOR.VEHICLE.4,dat2$CONTRIBUTING.FACTOR.VEHICLE.5)
#adding another column for no of vehicles involved in collision 
data2016$veh_inv = 5-(rowSums(is.na(data2016[2:6]) | data2016[2:6] == "" | data2016[2:6] == " "))
data2016 = data2016 %>% group_by(dat2.ZIP.CODE) %>% summarise(veh_inv = sum(veh_inv))
data2016= data2016[order((data2016$veh_inv),decreasing=TRUE),]
#removing row of null values
data2016 = data2016[-1,]
data2016[1,2]
```


#5243 for zip code 11207 being the highest . Ans Q5
```{r}

#splitting data from year 2013-2018
data13_18=data[data$DATE <= "2018-12-31" & data$DATE >= "2013-01-01",]
data13_18$year <- format(as.Date(data13_18$DATE), "%Y")
data3 <-data.frame(data13_18$year)
year13_18_col<-count(data3,data3$data13_18.year)
colnames(year13_18_col)[1] = "year"
colnames(year13_18_col)[2] = "NoCol"
year13_18_col = within(year13_18_col, { year <- as.numeric(as.character(year))})
#performing linear regression
model = lm(NoCol~year,data=year13_18_col)
slope = model$coef
slope[2]
```

#6448.171429 slope value. Q6

```{r}
#splitting data for year 2017 alone
data2017_d = data[data$DATE <= "2017-12-31"& data$DATE>= "2017-01-01",]
data2017_d$month = format(as.Date(data2017_d$DATE), "%m")
data2017 = data.frame(data2017_d$month,data2017_d$CONTRIBUTING.FACTOR.VEHICLE.1,data2017_d$CONTRIBUTING.FACTOR.VEHICLE.2,data2017_d$CONTRIBUTING.FACTOR.VEHICLE.3,data2017_d$CONTRIBUTING.FACTOR.VEHICLE.4,data2017$CONTRIBUTING.FACTOR.VEHICLE.5)
data2017$veh_inv = 5-(rowSums(is.na(data2017) | data2017 == "" | data2017 == " "))
#making a copy
colnames(data2017)[1] = "month"
data2017$month <- as.numeric(data2017$month)
data2017_veh3 = data2017
data2017_veh3 = subset(data2017_veh3, veh_inv>=3)
#removed data with vehicles involved less than 3
count_month_veh3 = count(data2017_veh3,data2017_veh3$month)
count_month_all = count(data2017,data2017$month)
prop_veh <- count_month_veh3$n / count_month_all$n
prop_veh
#gives an array of the prop of vehicles >3 involved in collision
null_count = sum(count_month_veh3$n)
exp_value = null_count/12
#1210 is the mean of multi collisions each month. calculated for chi squared
chi = cbind(count_month_veh3$n,exp_value)
chi_2 = rbind(chi[1],chi[4])
chi_3 = rbind(chi[1,2],chi[4,2])
chisq = cbind(chi_2,chi_3)
#table includes values of Observed collisions for month of jan and may as well as expected value that is mean 1210

chi_out = chisq.test(chisq)
chi_out$statistic
```

#0.2865976938 Chi Squared value . Ans Q7 

